Scalar_M1

### 2015-05-15 models/Model_Playspace.Rmd
model revision

```{r}
#cost function in length of chars
u.cost = function(u) {
  return(nchar(u))
}

#speaker likelihood helper 
#speaker.lhd = function(utt, info, alpha) {
#  exp(alpha*(log(info) - 3)) #Rething cost??? 
#}
speaker.lhd = function(d) {
  exp(4*(log(d) - 3)) #Rethink alpha and costs?
}

#speaker likelihood normalized
speaker.prob = function(rating, degree, m) {
  num = speaker.lhd(m[rating, degree])
  norm = 0
  for ( i in 1:2 ) {
    norm = norm + speaker.lhd(m[rating, i])
  }
  return(num / norm)
}

#non-normalized posterior
nn.post = function(rating, degree, m) {
  return(speaker.prob(rating, degree, m) * priors[rating, "prior.p"])
}

#normalized posterior
norm.post = function(rating, degree, m) {
  nn = nn.post(rating, degree, m)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(i, degree, m)
  }
  return(nn / norm)
}  
```

@@

```{r}
run.rsa <- function(d) {
  #BP------->
  #Get current matrix
  get.m = function(s, m) {
    m2 <- filter(m, scale == s) %>%
      spread(degree, speaker.p) %>%
      mutate(hi = hi / sum(hi), 
                  low = low / sum(low)) %>%
      select(hi, low)
    return(m2)
  }

  scale.matrix <- data %>% 
    select(stars, degree, speaker.p, prior.p) %>%
    mutate(listener.pred = norm.post(stars, degree, get.m(scale, .)))

  return scale.matrx$
 #BP------->

  #Old Mike code------>
  scale.matrix <- data %>% 
    select(stars, degree, speaker.p, prior.p) %>% 
    spread(degree, speaker.p) %>% 
    mutate(hi = hi / sum(hi), 
           low = low / sum(low)) %>%
    as.matrix

  preds <- expand.grid(stars = 1:5, 
              degree = c("low","hi")) %>%

    scale.matrix %>% 
    mutate(listener.pred = norm.post(stars, degree, get.m(scale, scale.matrix)))
  return(preds)
} 
```

@@

```{r}
data <- left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale) %>%
  do(model.preds = run.rsa(.))

```

>>>>

```{r}
#cost function in length of chars
# u.cost = function(u) {
#   return(nchar(u))
# }

#speaker likelihood helper 
speaker.lhd = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost)) #Rethink alpha and costs?
}

#speaker likelihood normalized
speaker.prob = function(rating, degree, m, alpha) {
  num = speaker.lhd(m[rating, degree], alpha)
  norm = 0
  for (i in 1:ncol(m)) {
    norm = norm + speaker.lhd(m[rating, i], alpha)
  }
  return(num / norm)
}

#non-normalized posterior
nn.post = function(rating, degree, m, alpha, useprior) {

  if (useprior) {
    prior <- priors[rating, "prior.p"]
  } else {
    prior = .2 
  }

  speaker.prob(rating, degree, m, alpha) * prior
}

#normalized posterior
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(i, degree, m, alpha, useprior)
  }
  return(nn / norm)
}  
```

@@

```{r}
run.rsa <- function(d, alpha = 1, useprior = TRUE, usenone = FALSE) {

  mat <- d %>%
    select(stars, degree, speaker.p) %>%
    spread(degree, speaker.p) %>%
    mutate(hi = hi / sum(hi), 
           low = low / sum(low)) %>%
    select(hi, low)

  if (usenone) {
    mat$none <- c(1, 0, 0, 0, 0)
  }

  d$pred <- round(as.numeric(mapply(norm.post, 
                                    d$stars, 
                                    d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
} 
```

@@

```{r}
data <- left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

run it.

```{r}
md <- data %>%
  do(run.rsa(., alpha = 1, useprior = FALSE, usenone = TRUE))

qplot(stars, listener.p, col=degree, 
      data = md) + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 2)
```
